---
title: "Sentiment analysis and visualization with R"
author: "Anelia Aleksandrova Boyadzhieva"
date: "2022-12-28"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparing the data

Installing libraries

```{r libraries}
#Loading libraries 
library(tidyverse)
library(sentimentr)
library(ggplot2)
library(dplyr)
```

## Loading data by using read.csv() and saving it in the variables redditY and redditX

```{r data}
redditY <- read.csv("TrollY_dataset.csv") #reading data for TrollY subreddit and assigning it to the redditY variable

redditX <- read.csv("TrollX_dataset.csv") #reading data for TrollX subreddit and assigning it to the redditX variable
```

## Cleaning the data

```{r}
#fuction for cleaning data -> renaming and removing columns
names <- function (x){
    x <- rename(x, post_score = data.score)
    x <- rename(x, date = Date)
    x <- rename(x, post = data.title)
    x <- x[, -5] #removing image url by column index 
    x <- x[, -2] #removing utc by column index 
}

#running the function on both datasets and saving into variables
redditXclean <- names(redditX)
redditYclean <- names(redditY)
```

# Doing sentiment analysis with sentiment()

```{r}
#using sentiment() method to do sentiment analysis on column 'data.title' from dataset 'redditY'
sentimentDataY <- sentiment(redditYclean$`post`) #assigning the new dataset to the variable 'sentimentDataY'
view(sentimentDataY) #viewing the data wich is stored in the variable sentimentDataY
sentimentDataY <- na.omit(sentimentDataY)

#using sentiment() method to do sentiment analysis on column 'data.title' from dataset 'redditX'
sentimentDataX <- sentiment(redditXclean$`post`) #assigning the new dataset to the variable 'sentimentDataX'
view(sentimentDataX)
sentimentDataX <- na.omit(sentimentDataX)
```

# Visualizing the data

## Comparison of sentiments

```{r}
boxplot(sentimentDataY$sentiment, sentimentDataX$sentiment, #getting data from dataset 'sentimentDataY' and 'sentimentDataX' column 'sentiment' 
        main = "Comparison of sentiments for subreddits TrollX and TrollY",
        xlab = "Sentiment",
        at = c(1,2), #positioning the two boxplots
        names = c("TrollY", "TrollX"), #naming the two boxplots
        col = c("green","lightblue"), #colour 
        horizontal = TRUE, #positioning
        frame = F
        )
```

## Individual boxplots visualizing sentiments

```{r}
#Making a boxplot for TrollX
boxplot(sentimentDataX$sentiment, #getting data from dataset 'sentimentDataX' column 'sentiment' 
        main = "Boxplot for sentiments in TrollX subreddit", #name/headline
        xlab = "Sentiment",
        col = "lightblue", #colour
        horizontal = TRUE, #positioning 
        frame = F
        )
#Making a boxplot for TrollY
boxplot(sentimentDataY$sentiment, #getting data from dataset 'sentimentDataY' column 'sentiment' 
        main = "Boxplot for sentiments in TrollY subreddit", #name/headline
        xlab = "Sentiment",
        col = "green",  #colour
        horizontal = TRUE, #positioning 
        frame = F
        )
```

## Individual plots showing the correlation between word count and sentiment

Here you can see two plots for the subreddits TrollX and TrollY. The sentiment on the X axis and word cound on the Y axis.

```{r}
#Making ggplot and customizing color, names and theme
wordCountSentimentX <- ggplot(data = sentimentDataX, mapping = aes(x = sort(sentimentDataX$sentiment, decreasing = TRUE), y = sentimentDataX$word_count)) + 
  geom_point(color = "dodgerblue") +
  labs(x = "Sentiment", y = "Word count", title = "Connection between word count and sentiment in TrollX") +
  theme_light() 

#Customizing title 
wordCountSentimentX + theme(
  plot.title = element_text(color="black", size=18, face="bold")
)

```

```{r}
#Making ggplot and customizing color, names and theme
wordCountSentimentY <- ggplot(data = sentimentDataY, mapping = aes(x = sort(sentimentDataY$sentiment, decreasing = TRUE), y = sentimentDataY$word_count)) + 
  geom_point(color = "green3") +
  labs(x= "Sentiment", y= "Word count", title = "Connection between word count and sentiment in TrollY") +
  theme_light()

#Customizing title 
wordCountSentimentY + theme(
  plot.title = element_text(color="black", size=18, face="bold")
)
```

## Boxplot comparing word count for TrollX and TrollY

There was a clear difference between the sentiment score and word count for the two subreddits. The following boxplot shows the word count for the subreddits, where it is evident that TrollX(blue) users use fewer words in their posts, whereas TrollY (green) users use less. This says something about how the users express themselves in the different subreddits.

```{r}
boxplot(sentimentDataY$word_count, sentimentDataX$word_count,  #getting data from dataset 'sentimentDataY' and 'sentimentDataX' column 'word_count' 
        main = "Comparison of word count for subreddits TrollX and TrollY",
        xlab = "Word count",
        at = c(1,2), #positioning the two boxplots
        names = c("TrollY", "TrollX"), #naming the two boxplots
        col = c("green","lightblue"),
        horizontal = TRUE,
        frame = F
        )
```
