---
title: "2) Using rvest to scrape data of my interest "
author: "Anelia Boyadzhieva"
date: "updated `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Github

Link to github, where you can find the rmd and html: https://github.com/Digital-Methods-HASS/au671033_Aleksandrova_Anelia

# Goal

My goal is to scrape data about the top free medical iPhone and iPad apps in App Store in Denmark (from the website similarweb.com). I wanted to scrape data for both iPhone and iPad, in order to compare the results and be able to look at which applications users like to use on different devices. 

# Challenge
Getting the dataframe from similarweb to look semi clean, like the one at the bottom from worldometers.

# Solution

First, I installed these libraries:

```{r libraries, warning=FALSE, message=FALSE}
library(rvest)
library(dplyr)
library(tidyr)
```

Then I wrote a function which takes a website as an argument, it scrapes it and creates a dataframe.

```{r scrape the site}
scrape_medical_apps <- function(website){
    url <- read_html(website)
    device_table <- url %>% 
            html_nodes("table") %>%
            html_table()  
  device_table <- do.call(cbind,unlist(device_table, recursive = FALSE))
}
```
## Scrape the data for top iPhone medical apps 
I gave the webpage as an argument and saved it in the variable table_iphone
```{r scrape iphone}
table_iphone <- scrape_medical_apps("https://www.similarweb.com/apps/top/apple/store-rank/dk/medical/top-free/iphone/")
table_iphone %>% 
    head()
```
## Scrape the data for top iPad medical apps
I gave the webpage as an argument and saved it in the variable table_ipad
```{r scrape ipad}
table_ipad <- scrape_medical_apps("https://www.similarweb.com/apps/top/apple/store-rank/dk/medical/top-free/ipad/")
table_ipad %>% 
    head()
```

# Trying out the function on another website 
I wanted to try and run the function on another website to see if it runs properly, because the other dataframes looked a little weird 
```{r srape test}
table_population <- scrape_medical_apps("https://www.worldometers.info/world-population/population-by-country/")
table_population %>%
  head()
```